---
title: Activity Report
author: "Rusell Glynn"
date: "10/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Read in dataset and process data

```{r libraries}

library(dplyr)
library(ggplot2)
library(readr)
library(lubridate)
library(naniar)
# read in data

rawexdata <- read.csv('activity.csv')

# remove NAs and convert date to factor

exdata <- rawexdata %>% filter(steps != 'NA')
exdata$date <- as.factor(ymd(exdata$date))
#exdata$interval <- factor(exdata$interval)

```

### 2. Create histogram of the total number to steps taken per day

```{r histogram, warning= F}

stepsperday <- exdata %>% group_by(date) %>% summarize(total_steps = sum(steps))
ggplot(stepsperday, aes(total_steps)) +
    geom_histogram()

```

### 3. Mean and median number of steps per day

```{r }

stepsperday %>% summarize(mean = mean(total_steps),
                          median =  median(total_steps))

```

### 4. Time series plot of the average number of steps taken

```{r}
meaninterval <- exdata %>% group_by(interval) %>% summarize(mean = mean(steps))
str(meaninterval)
ggplot(meaninterval, aes(interval, mean)) +
    geom_point() +
    ggtitle('Average Number of Steps During Each Time Interval') +
    xlab('interval minutes') +
    ylab('interval steps')

```

### 5. The 5-minute interval that, on average, contains the maximum nummber of steps

```{r}

maxinterval <- meaninterval %>% filter(mean == max(mean))
maxinterval$interval

```

### Code to describe missing data

```{r}

missingdata <- apply(is.na(rawexdata), 2, sum)
missingdata

```
All of the missing data are located in the steps variable.

Looking at the missing data to determine which intervals have missing data, we create a missing variable summary report (missdata).

```{r}
missdata <- rawexdata %>% group_by(interval) %>% select(interval, steps) %>% miss_var_summary()
head(missdata)
unique(missdata$n_miss)

```
It looks like every interval has exactly 8 missing values in the steps variable.

We can look at the number of consecutive rows are missing steps data.

```{r }

rawexdata$date <- as.factor(ymd(rawexdata$date))
missdatarun <- rawexdata %>% group_by(interval) %>%
    select(interval, steps) %>% miss_var_run(steps)
sumrun <- missdatarun %>% summary
sumrun

boxplot(missdatarun$run_length)
```

We can see that the median missing run length is 2. However, we have two outliers at 23 and 15 with a median of 2.

### Devise a strategy for filling in all of the missing values in the dataset.

We will impute the missing values in the dataset by using the median or each interval. We will store the data in a new dataset called imputerawexdata.

```{r imputing}

# imputing by median of interval

navector <- which(is.na(rawexdata))

medianinterval <- rawexdata %>% group_by(interval) %>% summarise(median = median(steps, na.rm = T))
imputerawexdata <- rawexdata
getmedian <- function(x) {
    y <-which(medianinterval$interval == x)
    intsteps <- medianinterval[y,2]
}

for(i in seq_along(navector)) {
    y = navector[i]
    z <- rawexdata[y,3]
    q <- getmedian(as.numeric(z))
    imputerawexdata[y,1] <- q
}

```
### Create a histogram of imputed data total steps per day. Report the mean and median total steps per day using this new dataset.

```{r imputed}

stepsperday <- imputerawexdata %>% group_by(date) %>% summarize(total_steps = sum(steps))
ggplot(stepsperday, aes(total_steps)) +
    geom_histogram()

```

Mean and median of the new dataset:

```{r}
stepsperday %>% summarize(mean = mean(total_steps),
                          median =  median(total_steps))
```
The mean is considerably lower than the previously calculated mean with NAs excluded. The median is only slightly lower.

### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends.

```{r}
imputerawexdata$date <- as.Date(imputerawexdata$date)
imputerawexdata <- imputerawexdata %>% mutate(weekday = weekdays(date))
days <- imputerawexdata %>% group_by(weekday) %>% summarize(total_steps = sum(steps))
imputerawexdata <- imputerawexdata %>% 
    mutate(day_type = ifelse(weekday %in% c('Saturday', 'Sunday'),'weekend','weekday'))
imputerawexdata$day_type <- as.factor(imputerawexdata$day_type)
str(imputerawexdata)
imputerawexdata %>% ggplot(aes(interval, steps)) +
    geom_line() +
     facet_grid(day_type ~ .)


```


                                         ```